---
title: "Linear Regression"
author: "Darya Vanichkina"
exercises: 0
keypoints: somepoint
objectives: someobjective
questions: What is the meaning of life FIXME?
source: Rmd
start: 0
teaching: 30
bibliography: references.bib
---

First, let's load the required libraries. We will use the `caret` library for our ML tasks, and the `tidyverse` for general data processing and visualisation.

```{r loadLibraries}
library(tidyverse)
library(caret)
library(naniar) # for visualising missing data

```

We will use the Ames housing data to explore different ML approaches to regression. This dataset was designed by Dean De Cock [@de2011ames] as an alternative to the "classic" Boston housing dataset, and has been extensively used in  ML teaching. It is also available from kaggle as part of its [advanced regression practice competition](https://www.kaggle.com/c/house-prices-advanced-regression-techniques).

The Ames Housing Data Documentation file describes the independent variables presented in the data. This includes:
- 20 continuous variables relate to various area dimensions for each observation
- 14 discrete variables, which typically quantify the number of items occurring within the house
- 23 ordinal, 23 nominal categorical variables, with 2 (STREET: gravel or paved) - 28 (NEIGHBORHOOD) classes 


```{r loadData}
ameshousing <- read_csv("data/AmesHousing.csv", guess_max=1500)
```

## Exploratory data analysis

```{r dim}
dim(ameshousing)
```

```{r str}
str(ameshousing)
```


```{r missingdata}
colSums(sapply(ameshousing, is.na)) %>% 
  as.data.frame() %>% 
  rename(Missing = ".") %>%
  tibble::rownames_to_column()%>% 
  arrange(desc(Missing))
```

### Use the naniar library to visualise missing

Some new visualisation for missing data in the tidy context has been proposed [@tierney2018expanding]. See this [web page](http://naniar.njtierney.com/articles/naniar-visualisation.html) for more options for your own data.

```{r uninformativeMissing}
gg_miss_var(ameshousing)
```

```{r naniar}
gg_miss_upset(ameshousing)
```

## EDA of outcome variable

```{r salesPrice}
ameshousing %>% 
  select(SalePrice) %>%
  ggplot(aes(x = SalePrice)) + geom_histogram(bins = 50) + theme_minimal()
```

```{r ScatterFindOutliers}
ameshousing %>%
  select(SalePrice, `Gr Liv Area`) %>%
  ggplot(aes(x = `Gr Liv Area`, y = SalePrice)) + geom_point() + theme_minimal() +
  geom_smooth(method= "lm")

```

```{r RemoveOutliers}
# remove 5 observations
ameshousingFilt <- 
  ameshousing %>% 
  filter(`Gr Liv Area` <= 4000)
```

Explore the variance:

```{r varianceExplore}
ameshousing %>%
  mutate()

```





## Splitting into testing and training datasets

We will use t


FIXME: (a plot of SALE PRICE versus GR LIV AREA will quickly indicate these points). Three of them are true outliers (Partial Sales that likely don’t represent actual market values) and two of them are simply unusual sales (very large houses priced relatively appropriately). I would recommend removing any houses with more than 4000 square feet from the data set

FIXME: when we do train/test split, do so by ordering houses by price and then holding out the ith one.




```

## References


  
  